---
title: "week3_exercise"
author: "Wenke Zimmermann"
date: "2022-05-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Segmentation of Trajectories

## Load the necessary libraries ###
```{r message=FALSE, warning= FALSE}
library(tidyverse)    
library(lubridate)
library(SimilarityMeasures)
library(cowplot)
```

## Load Data Sabi
```{r}
sabi <- read_delim("sabi.csv",",")
```

## A) Specify a temporal windows v for in which to measure Euclidean distances
In the dataset, the sampling interval is 15 minutes. If we take a temporal window of 60 minutes, that would mean including 4 fixes. We need to calculate the following Euclidean distances (pos representing single location):

1) pos[n-2] to pos[n]
2) pos[n-1] to pos[n]
3) pos[n] to pos[n+1]
4) pos[n] to pos[n+2]


## B) Measure the distance from every point to every other point within this temporal window (v)
We use the formula for calculating the Euclidean distance in in combination with lead() and lag(). For example, to create the necessary offset of n-2, we use lag(x, 2). For each offset, we create one individual column.
```{r}
sabi <- sabi %>%
  mutate(
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),   # distance to pos -30 minutes
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),   # distance to pos -15 minutes
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), # distance to pos +15 mintues
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2)  # distance to pos +30 minutes
  )
```

Now we want to calculate the mean distance of nMinus2, nMinus1, nPlus1, nPlus2 for each row. Since we want the mean value per Row, we have to explicitly specify this before mutate() with the function rowwise(). To remove this rowwise-grouping, we end the operation with ungroup().

Note that for the first two positions, we cannot calculate a stepMean since there is no Position n-2 for these positions. This is also true for the last to positions (lacking a position n+2).
```{r}
sabi <- sabi %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus2, nMinus1,nPlus1,nPlus2))
  ) %>%
  ungroup() 
```


## C) Remove “static points”: These are points where the average distance is less than a given threshold. This segments the trajectory into subtrajectories
We can now determine if an animal is moving or not by specifying a threshold distance on stepMean. In our example, we use the mean value as a threshold: Positions with distances below this value are considered static.
```{r}
sabi <- sabi %>% 
  ungroup() %>%
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

sabi_filter <- sabi %>%
  filter(!static)

sabi_filter%>%
  ggplot(aes(E, N)) +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")
```


## D) Now remove short subtrajectories: These are trajectories with a short duration (whereas “short” is tbd)
```{r}
ggplot()+
  geom_path(aes(E,N), sabi_filter %>% filter(stepMean > 500))
#Grafik ist falsch, aber ohne weitere Fehlersuche mal zur richtigen Aufgabenstellung übergegangen
```



## Task 1:Segmentation
# Load Data
```{r}
caro <- read_delim("caro60.csv",",")
```

# Sampling Interval is 1 minute. Use temporal window v of 6 minutes, window size of 6 positions (n +- 3)
```{r}
caro <- caro %>%
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),    # distance to pos -3 minutes
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),    # distance to pos -2 minutes
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),    # distance to pos -1 minutes
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2),  # distance to pos +1 mintues
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2),  # distance to pos +2 minutes
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)   # distance to pos +3 minutes
  )
```



## Task 2: Specify and apply threshold d
#Calculating stepMean for every point to every other point within this temporal window (v)
```{r}
caro <- caro %>%
  rowwise() %>%
  mutate(stepMean = mean(c(nMinus3,nMinus2, nMinus1,nPlus1,nPlus2,nPlus3))) %>%
  ungroup()  
```

#Explore stepMean to find reasonable threshold
```{r}
summary(caro$stepMean)

ggplot(caro, aes(x=stepMean))+         
  geom_histogram(binwidth=2) #zeigt meiste Werte liegen bei 5-10
```

#Remove static values
```{r}
caro <- caro %>% 
# ungroup() %>%    #nicht ganz klar warum hier wieder entgruppiert wird
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))  #für 'steMean' keinen! gewählten Threshold einsetzen

```



## Task 3: Visualize segmented trajectories with Boolean
```{r}
ggplot(caro, aes(E, N, color=static))+
  geom_path()+
  geom_point()+
  coord_equal()
#Plot zeigt 'NA'-Werte als Ausreisser, die nicht logisch erscheinen
```



## Task 4: Segment-based analysis
# Assign unique ID for each segment, that we can use as grouping variable
```{r}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
  }
#rle_id ist Name der Funktion, alles danach ist das was diese tut
```

#Use function to assign unique IDs to subtrajectories.
```{r}
caro <- caro %>%
  mutate(segment_id = rle_id(static))
```

#Visualize moving segments by colourizing them by segment_ID.
```{r}
caro %>%
  filter(static == FALSE) %>%   #Filter, damit nur Segmente in denen sich etwas bewegt, gezeigt werden; im Beispiel stand "filter(!static)" --> umgeschrieben um doppelte Verneinung zu vermeiden 
  ggplot(aes(E, N, color=segment_id))+
  geom_path()+
  geom_point()+
  coord_equal()+
  labs(title = "All segments (uncleaned)")
```

#Use segment_ID as a grouping variable to determine the segments duration and remove short segments (e.g. segments with a duration < 5 Minutes)
```{r}
caro_cleaned <- caro %>%
  group_by(segment_id) %>%
  mutate(duration = max(DatetimeUTC)-min(DatetimeUTC))%>% #berechne Dauer der Bewegung
  filter(duration > 300)%>%  #entferne alle Segmente mit Bewegung < 5 Minuten
  filter(static == FALSE)   #entferne alle "unbeweglichen" Punkte (hätte man sicher auch schon oben machen können)

ggplot(caro_cleaned, aes(E, N, color = segment_id))+
  geom_path()+
  geom_point()+
  coord_equal()+
  labs(title = "Long segments (removed segments <5)")
#gleicht Grafik in Beispiel, aber hat noch mehr Segmente, vermutlich Threshold zu klein gewählt

```
  
  

## Task 5: Similarity measures on human movement
# Achtung neues Dataset
```{r}
pedestrian <- read_delim("pedestrian.csv",",")
```

#Explore the trajectories first and get an idea on how the pedestrians moved
```{r}
ggplot(pedestrian, aes(E, N, color = TrajID))+
  geom_path()+
  geom_point()+
  coord_equal()+
  theme(legend.position = "none")+
  labs(title = "Visual comparison of the 6 trajectories",
      subtitle = "Each subplot highlights a trajectory")+
  facet_wrap(~ TrajID)
```



## Task 6: Calculate similarity
#Install package SimilarityMeasures
```{r}
#install.packages("SimilarityMeasures")
#am Anfang Datei ausgeführt
```

#Get familiar with the package
```{r}
?SimilarityMeasures
```

#All functions in the package need matrices as input, with one trajectory per matrix.
```{r}
#Select and safe single trajectories
TJ1 <- pedestrian %>%
  filter(TrajID == 1)
TJ2 <- pedestrian %>%
  filter(TrajID == 2)
TJ3 <- pedestrian %>%
  filter(TrajID == 3)
TJ4 <- pedestrian %>%
  filter(TrajID == 4)
TJ5 <- pedestrian %>%
  filter(TrajID == 5)
TJ6 <- pedestrian %>%
  filter(TrajID == 6)

#Change from dataframe to matrix
TJ1_mtx <- data.matrix(TJ1)
TJ2_mtx <- data.matrix(TJ2)
TJ3_mtx <- data.matrix(TJ3)
TJ4_mtx <- data.matrix(TJ4)
TJ5_mtx <- data.matrix(TJ5)
TJ6_mtx <- data.matrix(TJ6)

```


# Calculate DTW (dynamic time warping)
```{r}
#DTW(traj1, traj2, pointSpacing=-1)
#point spacing: integer value of the maximum index difference between trajectory1 and trajectory2 allowed in the calculation. A negative value sets the point spacing to unlimited.
TJ2_dtw <- DTW(TJ1_mtx, TJ2_mtx, pointSpacing = 4)
TJ3_dtw <- DTW(TJ1_mtx, TJ3_mtx, pointSpacing = 4)
TJ4_dtw <- DTW(TJ1_mtx, TJ4_mtx, pointSpacing = 4)
TJ5_dtw <- DTW(TJ1_mtx, TJ5_mtx, pointSpacing = 4)
TJ6_dtw <- DTW(TJ1_mtx, TJ6_mtx, pointSpacing = 4)

```

#Calculate Edit Distance
```{r}
#EditDist(traj1, traj2, pointDistance=20)
#point distance: floating point number representing maximum distance in each dimension allowed for points to be considered equivalent.
TJ2_editdist <- EditDist(TJ1_mtx, TJ2_mtx, pointDistance = 4)
TJ3_editdist <- EditDist(TJ1_mtx, TJ3_mtx, pointDistance = 4)
TJ4_editdist <- EditDist(TJ1_mtx, TJ4_mtx, pointDistance = 4)
TJ5_editdist <- EditDist(TJ1_mtx, TJ5_mtx, pointDistance = 4)
TJ6_editdist <- EditDist(TJ1_mtx, TJ6_mtx, pointDistance = 4)

```

#Calculate Frechet
```{r}
#Frechet(traj1, traj2, testLeash=-1)
#testleash: numeric value, which if positive, checks whether the leash can be used --> results in boolean; if values negative, then it is not used and the standard calculation is performed.
TJ2_frechet <- Frechet(TJ1_mtx, TJ2_mtx, testLeash = -1)
TJ3_frechet <- Frechet(TJ1_mtx, TJ3_mtx, testLeash = -1)
TJ4_frechet <- Frechet(TJ1_mtx, TJ4_mtx, testLeash = -1)
TJ5_frechet <- Frechet(TJ1_mtx, TJ5_mtx, testLeash = -1)
TJ6_frechet <- Frechet(TJ1_mtx, TJ6_mtx, testLeash = -1)

```

#Calculate LCSS (allowing translations)
```{r}
#LCSS(traj1, traj2, pointSpacing=-1, pointDistance=20, errorMarg=2, returnTrans=FALSE)
#pointSpacing: integer value of the maximum index difference between trajectory1 and trajectory2 allowed in the calculation. A negative value sets the point spacing to unlimited.
# pointDistance: floating point number representing the maximum distance in each dimension allowed for points to be considered equivalent.
#errorMarg: floating point error margin used to scale the accuracy and speed of the calculation.
#returnTrans: boolean value to allow the best translation found to be returned as well as the LCSS value if set to true
TJ2_lcss <- LCSS(TJ1_mtx, TJ2_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ3_lcss <- LCSS(TJ1_mtx, TJ3_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ4_lcss <- LCSS(TJ1_mtx, TJ4_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ5_lcss <- LCSS(TJ1_mtx, TJ5_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)
TJ6_lcss <- LCSS(TJ1_mtx, TJ6_mtx, pointSpacing = 4, pointDistance = 4, errorMarg = 2, returnTrans = FALSE)

```

#Create Tibble/Dataframe from calculated values
```{r}
#Zusammenfügen aller errechneten Werte pro Measure Methode

#Option 1: Dataframe
#würde theoretische funktionieren, aber teilweise nicht nachvollziehbare Values und Fehlermeldungen
#Traj_all <- data.frame(
#       "dtw" = c(TJ2_dtw,TJ3_dtw,TJ4_dtw,TJ5_dtw,TJ6_dtw),          
#       "editdist"= c(TJ2_editdist,TJ3_editdist,TJ4_editdist,TJ5_editdist,TJ6_editdist),
#       "frechet" = c(TJ2_frechet,TJ3_frechet,TJ4_frechet,TJ5_frechet,TJ6_frechet),
#       "lcss" = c(TJ2_lcss,TJ3_lcss,TJ4_lcss,TJ5_lcss,TJ6_lcss),
#       "trajectory" = c(2,3,4,5,6))


#Option 2: Tibble
dtw <- c(TJ2_dtw, TJ3_dtw, TJ4_dtw, TJ5_dtw, TJ6_dtw)
editdist <- c(TJ2_editdist, TJ3_editdist, TJ4_editdist, TJ5_editdist, TJ6_editdist)
frechet <- c(TJ2_frechet, TJ3_frechet, TJ4_frechet, TJ5_frechet, TJ6_frechet)
lcss <- c(TJ2_lcss, TJ3_lcss, TJ4_lcss, TJ5_lcss, TJ6_lcss)

Traj_all <- tibble(dtw, editdist, frechet, lcss, trajectory = c(2,3,4,5,6))

#Umformen des Tibbles (oder Dataframes)
Traj_long <- Traj_all %>%
  pivot_longer(names_to = "name", values_to = "value", -trajectory)

```

#Visualize and compare Traj1 to Traj2-6 using different similarity measures
```{r}
#Prepare single plots for each measure method
dtw_plot <- Traj_long %>%
  filter(name == "dtw") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "DTW")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))

editdist_plot <- Traj_long %>%
  filter(name == "editdist") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "Edit Distance")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))

frechet_plot <- Traj_long %>%
  filter(name == "frechet") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "Frechet")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))

lcss_plot <- Traj_long %>%
  filter(name == "lcss") %>%
  ggplot()+
  theme(legend.position = "none")+
  labs(x="",title = "LCSS")+
  geom_col(mapping = aes(x = trajectory, y = value, fill = trajectory))

#Zusammenfügen aller Plots
plot_grid(dtw_plot, editdist_plot, frechet_plot, lcss_plot)
#könnte noch Legende und Überschrift vertragen
#Grafiken DTW und LCSS sehen anders aus als Beispiel

```




